---
title: "DATA1204 - Assignment6"
author: "Chinedu Onyeka"
date: "12/18/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Load Libraries**
```{r load-libraries, message=FALSE}
# load libraries
library(tidyverse)
```

**Read the file**
```{r read-file, message=FALSE}
# read the file
url <- "https://raw.githubusercontent.com/chinedu2301/DC_Analytics/main/MultiRegDataset.csv"
hughes_dataset <- read_csv(url)
```

**Check the head**
```{r head}
# check the head
head(hughes_dataset)
```

**Check the structure**
```{r glimpse}
# Check the datatype and structure of the dataset
glimpse(hughes_dataset)
```

*There are 7 columns (variables) and 1,338 rows (observations) in the dataset.*  

<br>
**Descriptive Statistics**
```{r statistics}
summary(hughes_dataset)
```

*Age*
```{r sd-age}
sd(hughes_dataset$age)
```

The minimum age is 18 years while the maximum age is 64. The mean age is about 39 years and the standard deviation of age is about 14 years.  

*bmi*  
```{r}
sd(hughes_dataset$bmi)
```

The minimum bmi is 16 while the maximum bmi is 53.10. The mean bmi is about 30.67 and the standard deviation of bmi is about 6.1.  

*children*  
```{r}
sd(hughes_dataset$children)
```
The minimum number of children is 0 while the maximum number of children is 5. The mean number of children is 1 and the standard deviation of number of children is 1.  

*sex*  
There are 676 Males and 662 females in the dataset.  

*smoker*  
There are 274 who are smokers and 1064 who are not in the dataset.  

**Histogram - expenses**  
```{r histogram}
# histogram
hist(hughes_dataset$expenses, main = "Histogram of Expenses", xlab = "Expenses", col = "yellow")
```


**T-test**  

*Hypothesis Testing*  
Null hypothesis $H_{0}: \mu = 10,000$  
Alternative hypothesis $H_{1}: \mu \neq 10,000$

t-test
```{r t-test}
# t-test
t.test(hughes_dataset$expenses, mu = 10000)
```

*The p-value is less than 0.05, so we reject the null hypothesis that the mean of expenses is equal to 10,000*  
<br>
**Simple Linear Regression**  
y = Expenses, x = Smoker  
Expenses = Intercept + slope * Smoker  

*Train the model*

```{r}
# train the model  
simple_model <- lm(expenses ~ smoker, data = hughes_dataset)  

# get the summary
summary(simple_model)
```

The "smoker-yes"is significant, and the R-squared is 61.98%  
The simple linear regression model is $expenses = 8434.3 + 23616*smokeryes$  


*Evaluate the model*  
```{r plot-models-simple}
par(mfrow = c(1, 2))
# plot the fitted vs residual
plot(fitted(simple_model), resid(simple_model), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

# Q-Q plot
qqnorm(resid(simple_model), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(simple_model), col = "dodgerblue", lwd = 2)
```
<br>
**Multiple Linear Regression**  
$yi = B0 + B1xi1 + B2xi2 + ... + Bpxip + E$ where i = 1,2, ..., n  

*Hypothesis Statement*  
$H_{0} : β1 = β2 = ... = βk = 0$
$H_{1}:$ at least one $βi ≠ 0, i = 1, ..., k$  


*Train the model*

```{r}
# train the model  
multiple_model <- lm(expenses ~ ., data = hughes_dataset)  

# get the summary
summary(multiple_model)
```

*The R-squared is 75.09% for the multi-linear regression*  


$expenses = -11941.6 + 256.8*age - 131.3*sexmale + 339.3*bmi + 475.7*children +$  
$23847.5*smokeryes - 352.8* regionnorthwest - 1035.6*regionsoutheast -959.3*regionsouthwest$  

*Evaluate the model*  
```{r plot-models-multiple}
par(mfrow = c(1, 2))
# plot the fitted vs residual
plot(fitted(multiple_model), resid(multiple_model), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)

# Q-Q plot
qqnorm(resid(multiple_model), main = "Normal Q-Q Plot", col = "darkgrey")
qqline(resid(multiple_model), col = "dodgerblue", lwd = 2)
```

**Conclusion**  
Smoking has a positive effect on the expenses for the simple linear regression. Also, other input variables helped in increasing the R-squared values for predicting the expenses.













