---
title: "DATA 1204 - Assignment5"
author: "Chinedu Onyeka"
date: "12/11/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Load Libraries**

```{r load-libraries, message=FALSE}
# Load libraries
library(tidyverse)
```


**Hypothesis Statement**  



$H_{0}: \beta = 0$, coefficient $\beta$ of the dividend is zero and not statistically significant  
$H_{1}: \beta \neq 0$, co-efficient $\beta$ of dividend is not equal to zero and is statistically significant  

**Read the file**  
```{r read-file, message=FALSE}
url <- "https://raw.githubusercontent.com/chinedu2301/DC_Analytics/main/ols_stock.csv"
stock <- read_csv(url)
```

**Check the head**  
```{r head}
# check the head
head(stock)
```
**Check summary of the data**  
```{r summary}
# check the summary
summary(stock)
```

```{r linear-regression}
# fit the model
linear_model <- lm(stock_return_scaled ~ dividend, data = stock) 

# check the summary of the model
summary(linear_model)
```

```{r get-coefficients}
# get the model co-efficients
summary(linear_model)$coefficient
```

**Insights**

<li> The p-value for the dividend predictor is about 0.757925 > 0.05 (level of significance), therefore, we do not reject the null hypothesis. i.e we accept the null hypothesis that the co-efficient $\beta$ of the dividend predictor is zero and not statistically significant </li>
<li> From the model summary and coefficients, we can see that the intercept is about 204.7576, while the slope of the model is -12.9732. Therefore, the equation of the model is $Y(Stock Return Scaled) = 204.7576 - 12.9732*X(dividend)$ </li> 
<li> It can be seen that the model shows a negative relationship between stock_return_scaled and dividend. </li> 
<li> Also, the R-Squared is about 0.2% which indicates a poor model.


```{r r-squared}
# get the R-squared

summary(linear_model)$r.squared
```

From the R-Squared value, we see that the R square is 0.00204062 (0.2%) which indicates a poor model.  

```{r predict}
# Try a prediction
newdata <- data.frame(dividend = 1000) # wrap the parameter
predict(linear_model, newdata) # apply predict
```

**Variables to Include to help increase accuracy of the model**  
Sometimes a single predictor may not be enough to predict the target variable. To improve accuracy, more features or variables may have to be added. In this case, I will add the marketcap, and earnings_ranking. We can run a model that includes those extra two features to see if the model accuracy will improve.

```{r add-two-variables}
# fit the model with two additional variables
linear_model_more_variables <- lm(stock_return_scaled ~ 
                                    dividend + marketcap + earnings_ranking, data = stock)

# get the summary of the new model
summary(linear_model_more_variables)
```

From the model summary, we can see that the R-Squared for the model has increased drastically from 0.2% to about 54.73%. Also, from the p-values, `marketcap` and `earnings_ranking` are statistically significant predictors of the dependent variable (stock_return_scaled)









